Giant Low Surface Brightness Galaxies As Merger Remnants

### 2015-04-28 glsb.tex
add new greene paper
put in bib file
fixed spell checking

Observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.  Even the Galaxy Zoo project classifies GLSB galaxies as spirals.

The disks in GLSB systems are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  Disks of this size are a challenge for typicall $\Lambda$CDM cosmology galaxy formation and exotic formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies xxx-expand this a bit.  In their model, a collision where a thingy hits thuswise, blah blah, sets off an expanding ring of star formation.  As this the star formation subsides and the stetllar population ages, it fades to form a disk that resembles in size and exponential profile observed GLSBs.  XXX--also tie in \citep{Mapelli12}.

@@

Points to cover in intro
\begin{itemize}
\item{GLSB galaxies are really big spirals}
\item{Typically classified as extreme spirals observaionally}
\item{but HST observations of Malin~1 show it is a regular galaxy with a distinct extended disk \citet{Barth07}}
\item{Models for GLSB formation}
  \begin{itemize}
  \item{regular growth (maybe like polar ring galaxies, or just the large end of spirals)}
  \item{faded collisional ring. \citet{Mapelli08, Mapelli12} }
  \item{high angular momentum accretion event. \citet{Pen06}}
  \end{itemize}
  \item{Discuss how, if GLSBs are merger remnents, they represent a unique possibility to trace the dark matter halo of elliptical galaxies.  }
\end{itemize}

@@

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin 2 and $\sim -0.68$ for UGC 6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Greene12}.  Other cites for elliptical gradients \citep{Kuntschner10}

@@

Outline for discussion section
\begin{itemize}
\item{We find the GLSB systems Malin~2 and UGC~6614 are observationally most consistent with being normal elliptical galaxies that have undergone a high angular momentum minor merger with a gas rich system.}
\item{recap photometric evidence--the centers look just like SDSS ellipticals.}
\item{recap star formation histories. Disk is dominated by old stars.}
\item{Discuss metallicity gradient. Hard to imagine how such a metal poor disk could form right next to a very metal rich core.
      \citet{Raskutti14} use VIRUS-P to look at stellar halo kinematics of massive ellipticals.  Lots of good stuff on fast/slow rotators.  Also a very nice plot showing the radial gradients in age and metallicity as measured from Lick indices.  They don't see any sharp transition like I do, all very smooth age and metal curves. }
\item{extended comparison to \citet{Kasparova14}}
  \begin{itemize}
    \item{Mass model of Malin~2 requires a huge isothermal core radius.}
    \item{They measure gas metallicities and find solar in the center and half-solar at 20-30 kpc.  While it is a fool's errand to compare metallicities between stars and gas, it is interesting that the gas metallicity gradient is much smaller than in the stars.  If the central stellar metallicity is indeed super-solar and the gas metallicity only solar, that implies there must have been low-metallicity gas accretion onto the system at some point.  Hmm, looking at DR12 stuff, the pipeline reducitons are fitting solar metallicity to the center, not sure I believe that.  }
      \item{They have a nice mass-model that becomes dark matter dominated at around r=20 kpc.  Could this be the normal shape of elliptical dark matter halos?}
  \end{itemize}

@@

%\bibliography{/astro/users/yoachim/Dropbox/Papers/Bib_files/big_jabref}
\bibliography{/Users/yoachim/Dropbox/Papers/Bib_files/big_jabref.bib}

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Shmitz Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA PR1 DM
% LocalWords:  2HE IFU UGC HSB S0 SAURON ellipticals SB0 M31 Malin2 Spitzer IC
% LocalWords:  kpc HIPASS CCD metallicity MANGA 0pt SDSS arcsec lensing 7m IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement DR10 fiducial WMAP7 BAO H0 redshift
% LocalWords:  Faber Mg2 redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed CDM Kasparova LSB Dutton Raskutti Sa Morganti
%%  LocalWords:  Emonts Weijmans Ibata McConnachie Schweizer Magain
%%  LocalWords:  Oosterloo baryonic Buote GALEX Salim Kelson outlier
%%  LocalWords:  Bruzual Ahn Willett SQL Chilingarian Saulder Swaters
%%  LocalWords:  Krajn GN Kuntschner situ indices Emsell Capp CasJobs

>>>>

Observers tend to classify GLSB systems as normal spirals.  \citet{Matthews01b} list Malin~2 as an Scd while \citet{Das13} concludes that GLSB systems are an extreme form of late type spiral galaxies.  Similarly, \citet{Kasparova14} claim that the large dark matter scale of Malin~2 can explain it's features.  Even the Galaxy Zoo project classifies GLSB galaxies as spirals.  While GLSB systems superficially resemble normal spiral galaxies, they are clearly extreme outliers when their scale lengths are compared to other galaxy samples, being around a factor of 2 larger than the largest typical spirals \citep{Dalcanton97}.

The disks in GLSB systems are incredibly large, with Malin~1 being detected to a radius of over 100 kpc \citep{Moore06}.  Disks of this size are a challenge for typical $\Lambda$CDM cosmology galaxy formation and exotic formation scenarios have been proposed for GLSB systems.  \citet{Mapelli08} form GLSB-like galaxies via the fading of collisional ring galaxies.  In their model, a collision sets off an expanding ring of star formation.  As this the star formation subsides and the stellar population ages, the ring galaxy fades over $\sim$100 Myr to resembles observed GLSBs.  XXX-one problem is that most GLSB's are relatively isolated, and thus there are no obvious candidates for the impactor systems that would have triggered the star formation burst.

@@

Points to cover in intro
\begin{itemize}
\item{GLSB galaxies are really big spirals}
\item{Typically classified as extreme spirals obsessionally}
\item{but HST observations of Malin~1 show it is a regular galaxy with a distinct extended disk \citet{Barth07}}
\item{Models for GLSB formation}
  \begin{itemize}
  \item{regular growth (maybe like polar ring galaxies, or just the large end of spirals)}
  \item{faded collisional ring. \citet{Mapelli08, Mapelli12} }
  \item{high angular momentum accretion event. \citet{Pen06}}
  \end{itemize}
  \item{Discuss how, if GLSBs are merger remnants, they represent a unique possibility to trace the dark matter halo of elliptical galaxies.  }
\end{itemize}

@@

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin 2 and $\sim -0.68$ for UGC 6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Greene12, Greene15}.  Other cites for elliptical gradients \citep{Kuntschner10}

@@

Outline for discussion section
\begin{itemize}
\item{We find the GLSB systems Malin~2 and UGC~6614 are observationally most consistent with being normal elliptical galaxies that have undergone a high angular momentum minor merger with a gas rich system.}
\item{recap photometric evidence--the centers look just like SDSS ellipticals.}
\item{recap star formation histories. Disk is dominated by old stars.}
\item{Discuss metallicity gradient. Hard to imagine how such a metal poor disk could form right next to a very metal rich core.
      \citet{Raskutti14} use VIRUS-P to look at stellar halo kinematics of massive ellipticals.  Lots of good stuff on fast/slow rotators.  Also a very nice plot showing the radial gradients in age and metallicity as measured from Lick indices.  They don't see any sharp transition like I do, all very smooth age and metal curves. }
\item{extended comparison to \citet{Kasparova14}}
  \begin{itemize}
    \item{Mass model of Malin~2 requires a huge isothermal core radius.}
    \item{They measure gas metallicities and find solar in the center and half-solar at 20-30 kpc.  While it is a fool's errand to compare metallicities between stars and gas, it is interesting that the gas metallicity gradient is much smaller than in the stars.  If the central stellar metallicity is indeed super-solar and the gas metallicity only solar, that implies there must have been low-metallicity gas accretion onto the system at some point.  Hmm, looking at DR12 stuff, the pipeline reductions are fitting solar metallicity to the center, not sure I believe that.  }
      \item{They have a nice mass-model that becomes dark matter dominated at around r=20 kpc.  Could this be the normal shape of elliptical dark matter halos?}
  \end{itemize}

@@

%\bibliography{/astro/users/yoachim/Dropbox/Papers/Bib_files/big_jabref}
%\bibliography{/Users/yoachim/Dropbox/Papers/Bib_files/big_jabref.bib}
\bibliography{big_jabref.bib}

@@

% LocalWords:  Debattista Horrocks Lancashire RCUK Malin Stanonik Spavone Lelli
% LocalWords:  Bothun GLSB Das Mapelli Weij Loebman SungWon Kwak Scd Sd
% LocalWords:  Schiminovich Iodice Belokurov Fairchild RMS pPXF Cappellari th
% LocalWords:  GANDALF Sarzi CosmoloPy Bernardi GLSBs DeBlok Verheijen Yoachim
% LocalWords:  SEDs Brookhaven Instituto Astrofisica Canarias Notre JINA DM
% LocalWords:  IFU UGC HSB SAURON ellipticals Spitzer IC
% LocalWords:  kpc HIPASS CCD metallicity MANGA SDSS arcsec lensing IRAF
% LocalWords:  Spectrograph spectrograph IDL longslit photometric FWHM SSP mis
% LocalWords:  dispersions Balmer decrement fiducia redshift
% LocalWords:  Faber redshifts NGC kinemetry photometry anisotropy Gyr SFH
% LocalWords:  metallicities colliders rotators et al MNRAS de datacube Ppxf
% LocalWords:  dashed CDM Kasparova LSB Dutton Raskutti Sa Morganti
%%  LocalWords:  Emonts Weijmans Ibata McConnachie Schweizer Magain
%%  LocalWords:  Oosterloo baryonic Buote GALEX Salim Kelson outlier
%%  LocalWords:  Bruzual Ahn Willett SQL Chilingarian Saulder Swaters
%%  LocalWords:  Krajn GN Kuntschner situ indices Emsell Capp CasJobs
%%  LocalWords:  Schmitz circ accrete jabref CMD
